---
title: "High-Rate Processing"
author: "Al Cooper"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Ranadu)
```

# Objective

The "KalmanFilter.R" script is designed for 1-Hz data. Although the
processing is written to accomodate 25-Hz, the studies were done with
low-rate files and the standard script will fail at the end because
it does not handle the "Time" dimension properly for 25-Hz.  Because
the corrections vary only slowly, it is not necessary to calculate them
at high-rate.

For these reasons, it is more efficient to process 25-Hz files as follows:

1. First process the corresponding 1-Hz file using the 
"KalmanFilter.R" script, which will produce a new netCDF file
with a "KF" identifier; e.g., "WCR-TESTtf01KF.nc".  
2. From the resulting file, determine the correction factors applied to
the 1-Hz measurements by finding differences like PITCHKF-PITCH. The
key factors to determine are the corrections to the "state vector", which
consists of the three-component position vector, the three-component 
velocity vector, and the three orientation angles. The result of the
error-state Kalman filter is a set of corrections to this 9-component
state vector as a function of time.   
3. Add some reference variables to the 25-Hz file: ROC and ZROC, representing
the rate-of-climb and the integrated altitude from ROC, which as argued
in the NCAR Tech Note are preferable measurements to those provided by 
the INS. These can be added by the chunk "ROC.R" in the KalmanFilter 
repository.  
4. Apply the same error-state corrections to the 25-Hz file. This does
not require recalculation of the corrections or any further processing.
Instead, it only requires application of the 1-Hz corrections to the 
corresponding 25-Hz state vector.  
5. Once the corrections are applied to the state vector, calculations like
those leading to the wind must be repeated to get final variables like
WDKF, WSKF, and WIKF.  
6. Add the new 25-Hz variables to a copy of the 25-Hz netCDF file.

One complicating factor in this approach is that the attitude-angle 
components of the error-state vector have been smoothed in the "l-frame"
and then transformed to the "a-frame" where they are normally defined.
That is done because the error components vary more smoothly in the l-frame
(with coordinates local-north, local-east, local-up), while changes in
flight heading mix the error components as observed in the a-frame, with
coordinates starboard, longitudinally forward, and upward relative to the
aircraft. The addition of corrections to the 25-Hz file involves interpolation
of the 1-Hz results, so to preserve the smoothness of the 1-Hz correction
this may be best done in the l-frame and then transformed to the a-frame.
However, this first-try implementation does not do these transformations
but instead works directly in the a-frame to apply the corrections.

The corrections produced by the Kalman filter are these:  
```
Cor[, 1] = LATKF - LAT
Cor[, 2] = LONKF - LON
Cor[, 3] = ALTKF - ZROC
Cor[, 4] = VEWKF - VEW
Cor[, 5] = VNSKF - VNS
Cor[, 6] = ROCKF - ROC
Cor[, 7] = PITCHKF - PITCH
Cor[, 8] = ROLLKF - ROLL
Cor[, 9] = THDGKF - THDG
```

Once these are extracted from the 1-Hz file, they need to be interpolated
to 25-Hz values corresponding to the variables in the 25-Hz file. Then
they can be applied to those 25-Hz variables to obtain the new variables
for addition to the modified netCDF file.  

```{r include=TRUE, echo = TRUE}
VarList <- c('LAT', 'LON', 'VEW', 'VNS', 'PITCH', 'ROLL', 'THDG',
             'LATKF', 'LONKF', 'ALTKF', 'VEWKF', 'VNSKF', 'ROCKF', 'ATX',
             'AKKF', 'SSKF', 'ADIFR', 'BDIFR', 'QCF', 'TASX', 'GGVSPD', 'VSPD',
             'GGVEW', 'GGVNS',
             'PITCHKF', 'ROLLKF', 'THDGKF', 'PSXC', 'GGLAT', 'GGALT', 'ACINS')
D1 <- getNetCDF(setFileName('WCR-TEST', 'tf02KF'), VarList)
Rate <- 1
D1$Grav <- Gravity (D1$LAT, D1$GGALT)
Cradeg <- pi/180
Data <- D1  # ROC uses data.frame Data
source('chunks/ROC.R')
Cor <- rep(0, nrow(D1)*9)
dim(Cor) <- c(nrow(D1), 9)
D1 <- Data
Cor[, 1] = with(D1, LATKF - LAT)
Cor[, 2] = with(D1, LONKF - LON)
Cor[, 3] = with(D1, ALTKF - ZROC)
Cor[, 4] = with(D1, VEWKF - VEW)
Cor[, 5] = with(D1, VNSKF - VNS)
Cor[, 6] = with(D1, ROCKF - ROC)
Cor[, 7] = with(D1, PITCHKF - PITCH)
Cor[, 8] = with(D1, ROLLKF - ROLL)
Cor[, 9] = with(D1, THDGKF - THDG)
rm(Data)
## Get the needed variables from the 25-Hz file
fname25 <- setFileName('WCR-TEST', 'tf02h')
FI <- DataFileInfo(fname25)
VarList <- VarList[-which(!(VarList %in% FI$Variables))]
D25 <- getNetCDF(fname25, VarList)
D25$Grav <- Gravity (D25$LAT, D25$GGALT)
Data <- D25
Rate <- 25
source('chunks/ROC.R')
D25 <- Data
rm(Data)
```

Here is the interpolation function and code to match time intervals in case the 1-Hz and 25-Hz files cover different times:

```{r interp, include = TRUE}
## interpolate the results over the full data.frame:
IntFilter <- function (X, inRate, outRate) {
  if (inRate == outRate) {return (X)}
  ratio <- as.integer(outRate/inRate)    ## expected to be an integer
  x <- 0:(length(X)-1)
  A <- stats::approx (x, X, n=length(X)*ratio-ratio+1)
  T <- A$y
  T <- signal::filter(signal::sgolay(4,75),T)
  ## now shift to match Rate:
  n <- as.integer (ratio / 2)
  NL = length(T)
  T <- c(rep(T[1],n), T, rep(T[NL],ratio-n-1))  ## OK, even or odd ratio
  return (T)
}
## find time overlap
SE1 <- getStartEnd(D1)
SE25 <- getStartEnd(D25)
nsec <- function(T) {
  as.integer(T %% 100 + (T %/% 100) * 60 + (T %/% 10000) * 3600)
}
if(SE1[1] > SE25[1]) {
  gap <- nsec(SE1[1]) - nsec(SE25[1])
  for (n in 1:gap) {
    Cor <- rbind(Cor[1,], Cor)
  }
} else if (SE1[1] < SE25[1]) {
  gap <- nsec(SE25[1]) - nsec(SE1[1])
  Cor <- Cor[(gap+1):nrow(Cor), ]
}
if(SE1[2] > SE25[2]) {
  gap <- nsec(SE1[2]) - nsec(SE25[2])
  Cor <- Cor[1:(nrow(Cor) - gap), ]
} else if (SE1[2] < SE25[2]) {
  gap <- nsec(SE25[2]) - nsec(SE1[2])
  for (n in 1:gap) {
    Cor <- rbind(Cor, Cor[nrow(Cor), ])
  }
}
C25 <- vector('numeric', nrow(D25) * 9)
dim (C25) <- c(nrow(D25), 9)
for (j in 1:9) {
  C25[, j] <- IntFilter (Cor[, j], 1, 25)
}
```

The C25 correction matrix now can be used to calculate corrected values of the
components of the state vector:

```{r newvars, include = TRUE}
D25$LATKF <- D25$LAT + C25[, 1]
D25$LONKF <- D25$LON + C25[, 2]
D25$ALTKF <- D25$ZROC + C25[, 3]
D25$VEWKF <- D25$VEW + C25[, 4]
D25$VNSKF <- D25$VNS + C25[, 5]
D25$ROCKF <- D25$ROC + C25[, 6]
D25$PITCHKF <- D25$PITCH + C25[, 7]
D25$ROLLKF  <- D25$ROLL + C25[, 8]
D25$THDGKF  <- D25$THDG + C25[, 9]
```

Next the new values of AKKP and SSKP are added, as in KalmanFilter.R:

```{r radome, include = TRUE}
Platform <- ifelse (grepl('130', attr(D1, 'Platform')), "C130", "GV")
cffn <- 19.70547
cff <- 21.481
cfs <- c(4.525341674, 19.933222011, -0.001960992)
if (grepl('130', attr(D1, 'Platform'))) { # C-130 WECAN values
  cff <- 10.3123
  cfs <- c(5.6885, 14.0452, -0.00461)
}
D25$QR <- D25$ADIFR / D25$QCF
D25$QR[D25$QCF < 20] <- NA
D25$QR[is.infinite(D25$QR)] <- NA
CutoffFreq <- 600 * Rate
D25$QRS <- zoo::na.approx (as.vector(D25$QR), maxgap=1000*Rate, na.rm = FALSE, rule=2)
D25$QRS[is.na(D25$QRS)] <- 0
D25$QRS <- signal::filtfilt (signal::butter (3, 2/CutoffFreq), D25$QRS)
D25$QRF <-  D25$QR - D25$QRS
D25$QCFS <- zoo::na.approx (as.vector(D25$QCF), maxgap=1000*Rate, na.rm = FALSE)
D25$QCFS[is.na(D25$QCFS)] <- 0
D25$QCFS <- signal::filtfilt (signal::butter (3, 2/CutoffFreq), D25$QCFS)
D25$AKKF <- cff * D25$QRF + cfs[1] + cfs[2] * D25$QRS + cfs[3] * D25$QCFS
D25$AKKF[D25$QCF < 10] <- NA
D25$SSKF <- 0.008 + 22.301 * D25$BDIFR / D25$QCF
if (grepl('130', attr(D25, 'Platform'))) { # C-130 WECAN values; use with heading offset +0.76
  D25$SSKF <- 0.85 + 12.6582 * D25$BDIFR / D25$QCF
}
D25$SSKF[D25$QCF < 10] <- NA
```

The wind calculation then proceeds as follows:

```{r getWind, include = TRUE}
## This follows "chunks/wind-calculation.R" from the KalmanFilter directory
DataW <- D25
DataW$ATTACK <- DataW$AKKF  ## These are the values used in WindProcessor()
DataW$SSLIP <- DataW$SSKF
DataN <- WindProcessor (DataW, AC = Platform)  ## This function is in Ranadu
D25$WICC <- DataN$WIN
D25$WDCC <- DataN$WDN
D25$WSCC <- DataN$WSN
DataW$PITCH <- D25$PITCHKF
DataW$ROLL <- D25$ROLLKF
DataW$THDG <- D25$THDGKF
DataW$VEW <- D25$VEWKF
DataW$VNS <- D25$VNSKF
DataW$GGVSPD <- D25$ROCKF
DataN <- WindProcessor(DataW, AC = Platform, CompF=FALSE)    ## suppress comp filter and GPS lever arm)
D25$WDKF <- DataN$WDN
D25$WSKF <- DataN$WSN
D25$WIKF <- DataN$WIN
## add longitudinal and lateral components analogous to UXC and VYC:
.hdg <- D25$THDGKF * Cradeg
.wd <- D25$WDKF * Cradeg + pi
D25$UXKF <- D25$WSKF * (sin(.hdg)*sin(.wd) + cos(.hdg)*cos(.wd))
.hdg <- .hdg - pi/2
D25$VYKF <- D25$WSKF * (sin(.hdg)*sin(.wd) + cos(.hdg)*cos(.wd))
DataW$GGVSPD <- D25[, 'GGVSPD']   ## std script has code to substitute for GGVSPD
DataN <- WindProcessor(DataW, AC = Platform, CompF=FALSE)    ## suppress comp filter and GPS lever arm)
D25$WIKFG <- DataN$WIN
```

Now the data.frame "D25" contains all the KF-adjusted 25-Hz variables.
The remaining task is to write those new variables to a new netCDF file,
with appropriate name ending in 'hKF.nc'. First create the new file:

```{r newCDF, include = TRUE}
## create-new-netcdf.R

fnew <- sub ('h.nc', 'hKF.nc', fname25)
## beware: overwrites without warning!!
Z <- file.copy (fname25, fnew, overwrite=TRUE)  ## BEWARE: overwrites without warning!!

# function to copy attributes from old variable (e.g., PITCH) to new one (e.g., PITCHKF)
copy_attributes <- function (atv, v, nfile) {
  for (i in 1:length(atv)) {
    aname <- names(atv[i])
    if (grepl ('name', aname)) {next}  # skips long and standard names
    if (grepl ('units', aname)) {next}
    if (grepl ('Dependencies', aname)) {next}
    if (grepl ('actual_range', aname)) {next}
    if (is.numeric (atv[[i]])) {
      ncatt_put (nfile, v, attname=aname, attval=as.numeric(atv[[i]]))
    } else {
      ncatt_put (nfile, v, attname=aname, attval=as.character (atv[[i]]))
    }
  }
}
```

```{r writeCDF, include = FALSE}
## modify-new-netcdf.R

## variables needed for attributes:

VarListRef <- VarList
FI <- DataFileInfo (fname25)

if (!('GGVSPD' %in% FI$Variables)) {
  if ('GGVSPDB' %in% FI$Variables) {
    VarList [which (VarList == 'GGVSPD')] <- 'GGVSPDB'
  } else if ('VSPD_A' %in% FI$Variables) {
    VarList [which (VarList == 'GGVSPD')] <- 'VSPD_A'
  } else if ('VSPD_G' %in% FI$Variables) {
    VarList [which (VarList == 'GGVSPD')] <- 'VSPD_G'
  } else {
    print ('ERROR: no VSPD variable found')
    exit()
  }
}
for (Var in VarList) {
  if (!(Var %in% FI$Variables)) {
    print (sprintf (' required variable %s not found in file %s; skipping...', Var, fname))
    exit()
  }
}

netCDFfile <- nc_open (fnew, write=TRUE) 
Rate <- 25
Dimensions <- attr (D25, "Dimensions")
Dim <- Dimensions[["Time"]]
if ("sps25" %in% names (Dimensions)) {
  Rate <- 25
  Dim <- list(Dimensions[["sps25"]], Dimensions[["Time"]])
}
if ("sps50" %in% names (Dimensions)) {
  Rate <- 50
  Dim <- list(Dimensions[["sps50"]], Dimensions[["Time"]])
}
DATT <- D25  ## save to ensure that attributes are preserved

## variables to add to the netCDF file:
VarNew <- c('LATKF', 'LONKF', 'ALTKF', 'VEWKF', 'VNSKF', 'ROCKF', 'PITCHKF', 'ROLLKF', 'THDGKF',
            'WDKF', 'WSKF', 'WIKF', 'WICC', 'WDCC', 'WSCC', 'UXKF', 'VYKF', 
            'AKKF', 'SSKF')
VarOld <- c('LAT', 'LON', 'ALT', 'VEW', 'VNS', 'VSPD', 'PITCH', 'ROLL', 'THDG', 
            'WD', 'WS', 'WI', 'WIC', 'WDC', 'WSC', 'UXC', 'VYC', 'AKRD', 'SSRD')
VarUnits <- c('degrees', 'degrees', 'm', 'm/s', 'm/s', 'm/s', 'degrees', 'degrees', 'degrees',
              'degrees', 'm/s', 'm/s', 'm/s', 'degrees', 'm/s', 'm/s', 'm/s',
              'degrees', 'degrees')
VarStdName <- c('latitude, KF', 'longitude, KF', 'altitude MSL, KF',
                 'eastward groundspeed, KF', 'northward groundspeed, KF', 'rate of climb, KF',
                 'pitch, KF', 'roll, KF', 'heading, KF',
                 'wind direction, KF', 'wind speed, KF', 'vertical wind, KF', 
                 'vertical wind, KF and GGVSPD', 'WIC recalc', 'WDC recalc', 'WSC recalc',
                 'longitudinal wind, KF', 'lateral wind, KF', 
                 'angle of attack, KF', 'sideslip, KF')
VarLongName <- c('INS latitude, Kalman-filter-corrected',
                'INS longitude, Kalman-filter-corrected',
                'INS altitude, Kalman-filter-corrected',
                'INS eastward ground speed, Kalman-filter-corrected',
                'INS northward ground speed, Kalman-filter-corrected',
                'INS rate of climb, Kalman-filter-corrected',
                'INS aircraft pitch angle, Kalman-filter-corrected',
                'INS aircraft roll angle, Kalman-filter-corrected',
                'INS aircraft true heading angle, Kalman-filter-corrected',
                'horizontal wind direction, Kalman-filter-corrected',
                'horizontal wind speed, Kalman-filter-corrected',
                'vertical wind speed, Kalman-filter-corrected',
                'original WIC, recalculated',
                'original WDC, recalculated',
                'original WSC, recalculated',
                'longitudinal component of the horizontal wind, Kalman-filter-corrected',
                'lateral component of the horizontal wind, Kalman-filter-corrected',
                'angle of attack, from Kalman-filter routine', 'sideslip, from Kalman-filter routine')

## create the new variables
varCDF <- list ()
for (i in 1:length(VarNew)) {
  print (sprintf ('new-netcdf %d%% done', as.integer(100*(i-1)/length(VarNew))))
  varCDF[[i]] <- ncvar_def (VarNew[i],  
                            units=VarUnits[i], 
                            dim=Dim, 
                            missval=as.single(-32767.), prec='float', 
                            longname=VarLongName[i])
  if (i == 1) {
    newfile <- ncvar_add (netCDFfile, varCDF[[i]])
  } else {
    newfile <- ncvar_add (newfile, varCDF[[i]])
  }
  ATV <- ncatt_get (netCDFfile, VarOld[i])
  copy_attributes (ATV, VarNew[i], newfile)
  ncatt_put (newfile, VarNew[i], attname="standard_name", 
             attval=VarStdName[i])
  ncvar_put (newfile, varCDF[[i]], D25[, VarNew[i]], count=c(25, nrow(D25)/25))
}
nc_close (newfile)

```